Developing Address Clustering Heuristics for Account-Based Blockchain Networks: An Analysis based on a Specific Address Set

### 2023-08-27 paper.tex
Update paper

Previous work in this field can be broadly categorized into methods developed for blockchains that use the unspent transaction output (UTXO) model, such as Bitcoin, and those that utilize the account-based model, like Ethereum and Polygon. 

@@

\subsection{Ethereum, Polygon and the EVM}
Ethereum and Polygon are both smart contract-based account-model blockchains. They share the same execution logic - the Ethereum Virtual Machine (EVM).
The EVM provides a standardized framework for contract execution, ensuring that smart contracts produce deterministic results across all nodes in the network \citep{GW:14}. The EVM is independent of the underlying blockchain protocol, allowing other blockchain implementations to adopt this standardized framework for contract execution.\newline
\textit{Polygon} is a so-called sidechain and aims to address the scalability limitations of Ethereum. The adoption of the EVM means that both blockchains share the same user address schemes. It also enables developers to deploy and execute Ethereum-based smart contracts on the Polygon network with minimal changes. \citep{matic_whitepaper}

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world, with well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix actively engaging in it \citep{metaverse-retailing2023}.\newline
Two main reasons make Decentraland an ideal platform for empirical research.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' location within the metaverse and Web3 address. Using this capability, \cite{metaverse-retailing2023} captured snapshots of avatar locations over a nine-month period and provided the collected addresses for this thesis.\newline
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g., wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the avatars' economic activity. \newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users can buy or trade LAND or other in-game collectibles.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected over 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and also imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.

@@

\subsection{Self-Authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The ``self authorization`` heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. Since all the addresses' transactions are available in our data set, we can easily employ this heuristic.

\subsection{Deposit Address Reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. These addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded (since the exchange must pay for gas fees), and the time lag between receiving and forwarding funds.
However, our data set only includes the transactions or token transfers associated with the addresses. The transactions between deposit and exchange address are absent. Consequently, this clustering heuristic is not applicable and would require significantly more effort for gathering the required data.

\subsection{Airdrop Multi Participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that ever called the Uniswap V1 or V2 contracts until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He found around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defined multi-participation when two airdrop recipients forward their tokens to a single address. Since we do not have a whole view of the airdrop distribution events, we are not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process takes a long time and is not guaranteed to cluster many addresses, since more recent airdrop events were mostly carried out with some kind of sybil-resistance. Nevertheless, to test this approach we searched for addresses that forwarded the UNI airdrop to another address. We found eight such transfers, but none involved two addresses sending funds to a third address.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/time-of-day-activity.png}
	\caption{Time-of-Day Activity Histogram and KDE for a Selected ENS Address Pair (Based on \cite{Beres2020})}
	\label{fig:ToD}
\end{figure}

@@

\subsection{LAND Transfers}
In addition to the existing clustering heuristics, we propose our heuristic based on transfers of high-value NFTs such as LAND. We argue that, given the high value of Decentraland's land parcels, there is a strong likelihood that addresses are controlled by the same entity when LAND is transferred without any compensation. Users might perform such transfers to rearrange their LAND holdings, e.g., by upgrading to a more secure wallet. \newline
The primary challenge of this approach is to detect LAND transfers without any compensation. To accomplish this, we identify token or native asset transfers within the same transaction as a LAND transfer. Upon manually examining various implementations of LAND trades on the most popular NFT marketplace contracts, we noticed a recurring pattern: the user address appears in both the ``from`` and ``to`` fields if a trade occurs. Figure \ref{fig:LAND_example} illustrates an example of this pattern. 
In the case of pure LAND transfers, the sending and receiving addresses are the same for all transfers within a transaction. However, this approach does not consider other LAND trading schemes, including off-chain compensation or non-atomic (within multiple transactions) trade settlement.

@@

All three methods demonstrated remarkable efficacy in locating the target address, which indicates their ability to find multiple addresses controlled by a single entity.
Role2Vec is the most effective method, with the target address, on average, found within the 28 nearest neighbors. In contrast, Diff2Vec and DeepWalk produced slightly higher averages of 46 and 62, respectively. Furthermore, Role2Vec found the target address within the ten nearest neighbors in 24 out of 34 cases (71\%). Diff2Vec is barely less effective than Role2Vec or DeepWalk, although it is still generally successful. \newline
Notably, two address pairs are absent within the network graph and, therefore, display no values. However, they have been included to demonstrate that the intra-set asset transfer approach may only sometimes be applicable to cluster all addresses of a predefined set.

@@

A total of 740 LAND transfers were recorded, involving 413 unique addresses. Of these transfers, none involved any form of compensation within the same transaction. Upon analysis, we were able to group these addresses into 161 distinct clusters. Figure \ref{fig:LAND_clusters} visualizes the clusters in a network graph. Notably, the majority of clusters consist of two or three addresses. However, one cluster stands out, containing a significantly higher count of 71 addresses. This prominent cluster has been marked in green for ease of identification. \newline
LAND transfers between Decentraland addresses can also be visualized as a token network graph, as shown in Figure \ref{fig:LAND_network}. In this graph, red edges signify simple transfers, whereas blue edges represent LAND transfers made as part of a trade. Notably, the 71-address cluster previously mentioned is visible on the top-right of the largest connected component. \newline
A few key observations can be drawn from this network graph. In line with our expectations, most LAND transfers involve some form of compensation. Moreover, certain nodes play a central role in LAND trading, indicated by their high connectivity within the network. Lastly, a recurring behavior pattern is observed wherein an individual address purchases LAND and subsequently transfers it to a third address.

@@

\section{Conclusion}
This thesis aimed to determine how many unique entities correspond to the 473,927 addresses observed in Decentraland from July 16, 2022, to April 23, 2023. \newline
By focusing on Ethereum and Polygon on-chain data, we narrowed the set down to 137,544 addresses. The required data was efficiently sourced by gathering all transactions and token transfers related to the addresses via Etherscan and Polygonscan. Furthermore, with the help of ENS, we found a way to detect address pairs with a very high probability of being controlled by the same user. These address pairs served as our ground truth when evaluating the node embedding algorithms. \newline
We discussed several existing clustering heuristics and evaluated their suitability in our context. Among these heuristics, two were found to be appropriate for clustering a pre-defined set of addresses with the available data: self-authorization \citep{FV:17} and graph representation learning \citep{Beres2020}. We modified the graph representation learning approach for our research goal. %to align with our research goals, this is the first attempt to utilize intra-set asset transfers in constructing the network graph
To the best of our knowledge, we are the first to employ intra-set asset transfers for constructing the network graph. Moreover, we introduced a novel clustering heuristic that identifies uncompensated transfers of high-value NFTs, particularly LAND. \newline
Due to its general applicability, graph representation learning proved to be the most successful clustering heuristic, while the other two heuristics yielded far fewer address clusters. In our exploration of node embedding algorithms, the Role2Vec algorithm stood out, correctly identifying the ground truth pair within the ten nearest neighbors in 71 percent of the cases. \newline
From the 137,544 addresses transacting on Polygon or Ethereum, we identified an estimated 115,643 active Decentraland users. A subset of 8,483 entities, representing 7.3\%, operated through multiple addresses. Notably, the largest address cluster we observed was an entity with 500 distinct addresses. Based on our analysis, this entity is likely a bot targeting Decentraland NFT airdrops, with intentions to resell them later. \newline
To enhance clustering performance, node embeddings could incorporate time-of-day activity and gas price selection patterns. This approach holds considerable promise for future research in entity identification. \newline
A significant challenge for graph representation learning is the formation of clusters after node embedding. In our study, we relied on an Euclidean distance threshold for this purpose. However, the resulting number and size of clusters are highly sensitive to the precise selection of this threshold. \newline
Notably, our approach is versatile and can be applied to other large predefined address sets. This adaptability is particularly significant given the increasing importance of determining the true number of users in various protocols, especially in areas like Decentralized Finance and on-chain governance. \newline
Although we were able to detect entity clusters, Decentraland still exhibits a lot of natural activity. Its open infrastructure presents vast opportunities for in-depth analysis of the virtual world, as demonstrated in this thesis. \newline
Address clustering will continue to make further advancements. From a user perspective, it is crucial that privacy-enhancing protocols continue to evolve and gain large-scale adoption. Additionally, education on preserving privacy when transacting on blockchains becomes an increasingly important.

>>>>

Previous work in this field can be broadly categorized into methods developed for blockchains that use the unspent transaction output (UTXO) model, such as Bitcoin, and those that utilize the account-based model, like Ethereum and Polygon PoS. 

@@

\subsection{Ethereum, Polygon PoS and the EVM}
Ethereum and Polygon PoS are both smart contract-based account-model blockchains. They share the same execution logic - the Ethereum Virtual Machine (EVM).
The EVM provides a standardized framework for contract execution, ensuring that smart contracts produce deterministic results across all nodes in the network \citep{GW:14}. The EVM is independent of the underlying blockchain protocol, allowing other blockchain implementations to adopt this standardized framework for contract execution.\newline
\textit{Polygon PoS} (referred to as \textit{Polygon}) is a so-called sidechain and aims to address the scalability limitations of Ethereum. The adoption of the EVM means that both blockchains share the same user address schemes. It also enables developers to deploy and execute Ethereum-based smart contracts on the Polygon network with minimal changes. \citep{matic_whitepaper}

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world, with well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix actively engaging in it \citep{metaverse-retailing2023}.\newline
Two main reasons make Decentraland an ideal platform for empirical research.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' location within the metaverse and Web3 address. Using this capability, \cite{metaverse-retailing2023} captured snapshots of avatar locations over a nine-month period and provided the collected addresses for this thesis.\newline
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g., wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the avatars' economic activity. \newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, a NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users can buy or trade LAND or other in-game collectibles.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected over 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.

@@

\subsection{Self-Authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The \textit{self authorization} heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. Since all the addresses' transactions are available in our data set, we can easily employ this heuristic.

\subsection{Deposit Address Reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. These addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded, as the exchange must pay for gas fees, and the time lag between receiving and forwarding funds.
However, our data set only includes the transactions or token transfers associated with the addresses. The transactions between deposit and exchange address are absent. Consequently, this clustering heuristic is not applicable and would require significantly more effort for gathering the required data.

\subsection{Airdrop Multi Participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that ever called the Uniswap V1 or V2 contracts until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He found around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defined multi-participation when two airdrop recipients forward their tokens to a single address. Since we do not have a whole view of the airdrop distribution events, we are not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process takes a long time and is not guaranteed to cluster many addresses, since more recent airdrop events were mostly carried out with some kind of Sybil-resistance. Nevertheless, to test this approach we searched for addresses that forwarded the UNI airdrop to another address. We found eight such transfers, but none involved two addresses sending funds to a third address.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/time-of-day-activity.png}
	\caption{Time-of-Day Activity Histogram and KDE for a Selected ENS Address Pair, Own Illustration Based on \cite{Beres2020}}
	\label{fig:ToD}
\end{figure}

@@

\subsection{LAND Transfers}
In addition to the existing clustering heuristics, we propose our heuristic based on transfers of high-value NFTs such as LAND. We argue that, given the high value of Decentraland's land parcels, there is a strong likelihood that addresses are controlled by the same entity when LAND is transferred without any compensation. Users might perform such transfers to rearrange their LAND holdings, e.g., by upgrading to a more secure wallet. \newline
The primary challenge of this approach is to detect LAND transfers without any compensation. To accomplish this, we identify token or native asset transfers within the same transaction as a LAND transfer. Upon manually examining various implementations of LAND trades on the most popular NFT marketplace contracts, we noticed a recurring pattern: the user address appears in both the \texttt{``from``} and \texttt{``to``} fields if a trade occurs. Figure \ref{fig:LAND_example} illustrates an example of this pattern. 
In the case of pure LAND transfers, the sending and receiving addresses are the same for all transfers within a transaction. However, this approach does not consider other LAND trading schemes, including off-chain compensation or non-atomic (within multiple transactions) trade settlement.

@@

All three methods demonstrated remarkable efficacy in locating the target address, which indicates their ability to find multiple addresses controlled by a single entity.
Role2Vec is the most effective method, with the target address, on average, found within the 28 nearest neighbors. In contrast, Diff2Vec and DeepWalk produced slightly higher averages of 46 and 62, respectively. Furthermore, Role2Vec found the target address within the ten nearest neighbors in 24 out of 34 cases (71\%). Diff2Vec is barely less effective than Role2Vec or DeepWalk, although it is still generally successful. \newline
Notably, two address pairs are absent within the network graph and, therefore, display no values. However, they have been included to demonstrate that the intra-set asset transfer approach may not always be applicable to cluster all addresses of a predefined set.

@@

A total of 740 LAND transfers were recorded, involving 413 unique addresses. Of these transfers, none involved any form of compensation within the same transaction. Upon analysis, we were able to group these addresses into 161 distinct clusters. Figure \ref{fig:LAND_clusters} visualizes the clusters in a network graph. Notably, the majority of clusters consist of two or three addresses. However, one cluster stands out, containing a significantly higher count of 71 addresses. This prominent cluster has been marked in green for ease of identification. \newline
LAND transfers between Decentraland addresses can also be visualized as a token network graph, as shown in Figure \ref{fig:LAND_network}. In this graph, red edges signify simple transfers, whereas blue edges represent LAND transfers made as part of a trade. Notably, the 71-address cluster previously mentioned is visible on the top-right of the largest connected component. \newline
A few key observations can be drawn from this network graph. In line with our expectations, most LAND transfers involve some form of compensation. Moreover, certain nodes play a central role in LAND trading activity, indicated by their high connectivity within the network. Lastly, a recurring behavior pattern is observed wherein an individual address purchases LAND and subsequently transfers it to a third address.

@@

\section{Conclusion}
This thesis aimed to determine how many unique entities correspond to the 473,927 addresses observed in Decentraland from July 16, 2022, to April 23, 2023. \newline
By focusing on Ethereum and Polygon on-chain data, we narrowed the set down to 137,544 addresses. The required data was efficiently sourced by gathering all transactions and token transfers related to the addresses via Etherscan and Polygonscan. Furthermore, with the help of ENS, we found a way to detect address pairs with a very high probability of being controlled by the same user. These address pairs served as our ground truth when evaluating the node embedding algorithms. \newline
We discussed several existing clustering heuristics and evaluated their suitability in our context. Among these heuristics, two were found to be appropriate for clustering a predefined set of addresses with the available data: self-authorization \citep{FV:17} and graph representation learning \citep{Beres2020}. We modified the graph representation learning approach for our research goal. %to align with our research goals, this is the first attempt to utilize intra-set asset transfers in constructing the network graph
To the best of our knowledge, we are the first to employ intra-set asset transfers for constructing the network graph. Moreover, we introduced a novel clustering heuristic that identifies uncompensated transfers of high-value NFTs, particularly LAND. \newline
Due to its general applicability, graph representation learning proved to be the most successful clustering heuristic, while the other two heuristics yielded far fewer address clusters. In our exploration of node embedding algorithms, the Role2Vec algorithm stood out, correctly identifying the ground truth pair within the ten nearest neighbors in 71 percent of the cases. \newline
From the 137,544 addresses transacting on Polygon or Ethereum, we identified an estimated 115,643 active Decentraland users. A subset of 8,483 entities, representing 7.3\%, operated through multiple addresses. Notably, the largest address cluster we observed was an entity with 500 distinct addresses. Based on our analysis, this entity is likely a bot targeting Decentraland NFT airdrops, with intentions to resell them later. \newline
To enhance clustering performance, node embeddings could incorporate time-of-day activity and gas price selection patterns. This approach holds considerable promise for future research in entity identification. \newline
A significant challenge for graph representation learning is the formation of clusters after node embedding. In our study, we relied on an Euclidean distance threshold for this purpose. However, the resulting number and size of clusters are highly sensitive to the precise selection of this threshold. \newline
Notably, our approach is versatile and can be applied to other large predefined address sets. This adaptability is particularly significant given the increasing importance of determining the true number of users in various blockchain applications, especially in areas like Decentralized Finance or on-chain governance. \newline
Although we were able to detect entity clusters, Decentraland still exhibits a lot of natural activity. Its open infrastructure presents vast opportunities for in-depth analysis of the virtual world, as demonstrated in this thesis. \newline
Address clustering will continue to make further advancements. From a user perspective, it is crucial that privacy-enhancing protocols continue to evolve and gain large-scale adoption. Additionally, education on preserving privacy when transacting on blockchains becomes an increasingly important.
